---
title: "Summer Internship"
author: "Navin Ramanan"
date: "6/01/2022"
output: html_document
---

##Setup
```{r Packages + setup, include=FALSE, results='hide'}
#Source your main
source('APP_Sex.R') # load all our wonderful functions
knitr::opts_chunk$set(echo = FALSE) # no markdown for you <3

```

1.  Map Ensembl ID  to Gene Symbol in the counts file

```{r Part1}
#Read in verse_counts.tsv and display the head of the resulting tibble
Hippocampus_counts_filename2 <- '../data/Hippocampus_counts_matrix2.csv'
Cortex_counts_filename2 <- '../data/Cortex_counts_matrix2.csv'

Hippocampus_counts2 <- read.csv(Hippocampus_counts_filename2,as.is=TRUE, check.names=FALSE)
Cortex_counts2 <- read.csv(Cortex_counts_filename2,as.is=TRUE, check.names=FALSE)

Hippocampus_gene_name2 <- ensembl_id_to_gene_symbol(Hippocampus_counts2$Geneid)
Cortex_gene_name2 <- ensembl_id_to_gene_symbol(Cortex_counts2$Geneid)

```

2.  Write Counts Data with Gene Symbol to CSV

```{r Part2}

Hippocampus_geneids <- Hippocampus_gene_name2
Cortex_geneids <- Cortex_gene_name2
names(Hippocampus_geneids)[names(Hippocampus_geneids) == "ensembl_gene_id"] <- "Geneid"
names(Cortex_geneids)[names(Cortex_geneids) == "ensembl_gene_id"] <- "Geneid"

#remove duplicate rows across specific columns of data frame
Hippocampus_Geneid <- Hippocampus_geneids %>% distinct(Geneid, .keep_all = TRUE)
Cortex_Geneid <- Cortex_geneids %>% distinct(Geneid, .keep_all = TRUE)

new_Hippocampus_counts <- merge(Hippocampus_geneids, Hippocampus_counts2, by='Geneid')
new_Cortex_counts <- merge(Cortex_geneids, Cortex_counts2, by='Geneid')

# write counts table with mice gene name to a csv file
write.csv(new_Hippocampus_counts,'Hippocampus_counts_with_gene_names.csv')
write.csv(new_Cortex_counts,'Cortex_counts_with_gene_names.csv')

```

3.  Reading and filtering Hippocampus count and metadata files

```{r Part3}
#Read in verse_counts.tsv and display the head of the resulting tibble
Hippocampus_counts_filename <- '../data/Hippocampus_counts_matrix.csv'
Hippocampus_metadata_filename <- '../data/Hippocampus_metadata.csv'

# metadata
Hippocampus_metadata <- read.csv(Hippocampus_metadata_filename,as.is=TRUE)
# Make the first col of metadata into rownames
Hippocampus_meta_subset <- subset(Hippocampus_metadata, select= -c(1))
rownames(Hippocampus_meta_subset) <- Hippocampus_metadata[,1]

# Count data
Hippocampus_counts <- read.csv(Hippocampus_counts_filename,as.is=TRUE, check.names=FALSE)
# Keep only Counts cols and make Geneid col into rownames
Hippocampus_counts_subset <- subset(Hippocampus_counts, select= -c(1,2,3,4,5,6))
rownames(Hippocampus_counts_subset) <- Hippocampus_counts[,1]

# Keep rows with sum of sample counts greater than 10
Hippocampus_counts_subset_df <- Hippocampus_counts_subset[(rowSums(Hippocampus_counts_subset) > 10),]

```

4.  Reading and filtering Cortex count and metadata files

```{r Part4}
#Read in verse_counts.tsv and display the head of the resulting tibble
Cortex_counts_filename <- '../data/Cortex_counts_matrix.csv'
Cortex_metadata_filename <- '../data/Cortex_metadata.csv'

# metadata
Cortex_metadata <- read.csv(Cortex_metadata_filename,as.is=TRUE)
# Make the first col of metadata into rownames
Cortex_meta_subset <- subset(Cortex_metadata, select= -c(1))
rownames(Cortex_meta_subset) <- Cortex_metadata[,1]

# Count data
Cortex_counts <- read.csv(Cortex_counts_filename,as.is=TRUE, check.names=FALSE)
# Keep only Counts cols and make Geneid col into rownames
Cortex_counts_subset <- subset(Cortex_counts, select= -c(1,2,3,4,5,6))
rownames(Cortex_counts_subset) <- Cortex_counts[,1]

# Keep rows with sum of sample counts greater than 10
Cortex_counts_subset_df <- Cortex_counts_subset[(rowSums(Cortex_counts_subset) > 10),]

```

5.  Cortex: APP (APP vs WT) while controlling for covariate Sex (~ Sex + APP)
   
```{r Part5}

message("Cortex: Running DESeq2 for APP (APP vs WT) while controlling for covariate Sex (~ Sex + APP):padj < 0.05 and 0.10 ")
condition = "WT"
Cortex_deseq_results <- run_deseq_APP_Sex(Cortex_counts_subset_df, Cortex_meta_subset, 10, condition)
Cortex_results <- Cortex_deseq_results$res
Cortex_dds <- Cortex_deseq_results$dds
head(Cortex_results, 10)
Cortex_results
Cortex_dds
# write out DE results
write.csv(Cortex_results,'Cortex_deseq_results_APP_Sex.csv')
message("Cortex DESeq2 result is saved in file: 'Cortex_deseq_results_APP_Sex.csv' ")

# write out matrix of normalized counts
write.csv(counts(Cortex_dds,normalized=TRUE),'Cortex_deseq_norm_counts_APP_Sex.csv')
message("Cortex DESeq2 normalized counts is saved in file: 'Cortex_deseq_norm_counts_APP_Sex.csv' ")

#Add gene name to deseq results file and save the file
Cortex_deseq_res <- as.data.frame(Cortex_deseq_results$res)
Cortex_deseq_res_with_genename <- tibble::rownames_to_column(Cortex_deseq_res, "Geneid")
Cortex_deseq_res_with_genename$mgi_symbol <- new_Cortex_counts$mgi_symbol[match(Cortex_deseq_res_with_genename$Geneid, new_Cortex_counts$Geneid)]
Cortex_deseq_res_with_genename <- Cortex_deseq_res_with_genename %>% relocate(mgi_symbol, .after=Geneid)
write.csv(Cortex_deseq_res_with_genename, 'Cortex_deseq_results_with_genename_APP_Sex.csv')
message("Cortex DESeq2 result with Gene names (mgi_symbols) is saved in file: 'Cortex_deseq_results_with_genename_APP_Sex.csv' ")


############################## padj_threshold = 0.05 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.05
print(nrow(Cortex_deseq_res))
Cortex_deseq_results_padj_filtered <- filter(Cortex_deseq_res, padj < padj_threshold)
print(nrow(Cortex_deseq_results_padj_filtered))
write.csv(Cortex_deseq_results_padj_filtered, 'Cortex_deseq_results_APP_Sex_padj_05_filtered.csv')
message("Cortex DESeq2 result after padj (0.05) filtering is saved in file: 'Cortex_deseq_results_APP_Sex_padj_05_filtered.csv' ")

# Number of genes significant at padj less than 0.05
message(paste0("Number of DE genes significant at padj < 0.05 for Cortex:APP (APP vs WT) while controlling Sex: ", nrow(Cortex_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Cortex_deseq_res) - nrow(Cortex_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold 0.05 filtering for Cortex:APP(APP vs WT) while controlling Sex: ", no_of_genes_discarded))
  
#Histograms of fold change values from the significant DE genes. 
Cortex_deseq_hist_title1 <- "Histogram of log2FoldChange:Cortex:APP (APP vs WT): Sex Controlled:padj < 0.05"
Cortex_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_hist_title1)
Cortex_deseq_fc_hist_plot1

#Scatter plots of fold change vs nominal p-value.
Cortex_deseq_scat_title1 <- "Scatter Plot:Cortex:APP (APP vs WT) - Sex Controlled:padj < 0.05"
Cortex_deseq_scat_plot1 <- deseq_DE_scat_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_scat_title1)
Cortex_deseq_scat_plot1

#Add gene name to deseq results file and save the file
Cortex_deseq_padj_filtered_05 <- Cortex_deseq_results_padj_filtered
Cortex_deseq_padj_filtered_05 <- tibble::rownames_to_column(Cortex_deseq_padj_filtered_05, "Geneid")
Cortex_deseq_padj_filtered_05$mgi_symbol <- 
new_Cortex_counts$mgi_symbol[match(Cortex_deseq_padj_filtered_05$Geneid, new_Cortex_counts$Geneid)]
write.csv(Cortex_deseq_padj_filtered_05, 'Cortex_deseq_results_APP_Sex_padj_05_filtered_genename.csv')

############################## padj_threshold = 0.10 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.10
print(nrow(Cortex_deseq_res))
Cortex_deseq_results_padj_filtered <- filter(Cortex_deseq_res, padj < padj_threshold)
print(nrow(Cortex_deseq_results_padj_filtered))
write.csv(Cortex_deseq_results_padj_filtered, 'Cortex_deseq_results_APP_Sex_padj_1_filtered.csv')
message("Cortex DESeq2 result after padj (0.1) filtering is saved in file: 'Cortex_deseq_results_APP_Sex_padj_1_filtered.csv' ")

# Number of genes significant at padj less than 0.1
message(paste0("Number of DE genes significant at padj < 0.1 for Cortex:APP (APP vs WT)  while controlling Sex: ", nrow(Cortex_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Cortex_deseq_res) - nrow(Cortex_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold 0.1 filtering for Cortex:APP(APP vs WT)  while controlling Sex: ", no_of_genes_discarded))
  
#Histograms of fold change values from the significant DE genes. 
Cortex_deseq_hist_title1 <- "Histogram of log2FoldChange:Cortex:APP (APP vs WT) - Sex Controlled:padj < 0.1"
Cortex_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_hist_title1)
Cortex_deseq_fc_hist_plot1

#Scatter plots of fold change vs nominal p-value.
Cortex_deseq_scat_title1 <- "Scatter Plot:Cortex:APP (APP vs WT) - Sex Controlled:padj < 0.1"
Cortex_deseq_scat_plot1 <- deseq_DE_scat_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_scat_title1)
Cortex_deseq_scat_plot1

#Add gene name to deseq results file and save the file
Cortex_deseq_padj_filtered_1 <- Cortex_deseq_results_padj_filtered
Cortex_deseq_padj_filtered_1 <- tibble::rownames_to_column(Cortex_deseq_padj_filtered_1, "Geneid")
Cortex_deseq_padj_filtered_1$mgi_symbol <- new_Cortex_counts$mgi_symbol[match(Cortex_deseq_padj_filtered_1$Geneid, new_Cortex_counts$Geneid)]
write.csv(Cortex_deseq_padj_filtered_1, 'Cortex_deseq_results_APP_Sex_padj_1_filtered_genename.csv')

```

6. Cortex: Enhanced Volcano Plot

```{r Part6}
# generate Enhanced Volcano Plot
padj_threshold <- 0.05
FC_threshold <- 0
subtitle <- "DESeq2 Results:Cortex:APP (APP vs WT) - Sex Controlled"
Cortex_EnhancedVolcano_plot <- plot_enhanced_volcano(Cortex_deseq_res_with_genename, padj_threshold,  FC_threshold, subtitle)
Cortex_EnhancedVolcano_plot

```

7. Cortex: Heatmap

```{r Part7}
 # generate heatmap
 Cortex_title <- "Heatmap:Cortex:APP (APP vs WT) - Sex Controlled"
 Cortex_results_with_genename <- Cortex_deseq_res_with_genename[, -c(1)]
 rownames(Cortex_results_with_genename) <- Cortex_deseq_res_with_genename$Geneid
 plot_heatmap <-  generate_heatmap(Cortex_dds, Cortex_results_with_genename, Cortex_title)
 plot_heatmap

```

8 Cortex: PCA

```{r Part8}

 # PCA
 message("************** Principal Component Analysis (PCA) **************")
 Cortex_ntd <- normTransform(Cortex_dds)
 Cortex_ntd
 #view(assay(Cortex_ntd))
 Cortex_pca_results <- prcomp(scale(t(assay(Cortex_ntd))), center=FALSE, scale=FALSE)
 summary(Cortex_pca_results)
 Cortex_pc_variance_explained <- calculate_variance_explained(Cortex_pca_results)
 Cortex_pc_variance_explained
 Cortex_variance_tibble <- make_variance_tibble(Cortex_pc_variance_explained, Cortex_pca_results)
 head(Cortex_variance_tibble, 5)
 plot_pca_variance(Cortex_variance_tibble)
 Cortex_biplot_title = "PCA Plot: PC1 vs PC2:Cortex:APP (APP vs WT) - Sex Controlled"
 make_biplot(Cortex_metadata_filename, Cortex_pca_results, Cortex_biplot_title)
 
```

9.  Hippocampus: APP (APP vs WT) while controlling for covariate Sex (~ Sex + APP)

```{r Part9}
message("Hippocampus: Running DESeq2 for APP (APP vs WT) while controlling for covariate Sex (~ Sex + APP):padj < 0.05 and 0.10 ")
condition = "WT"
Hippocampus_deseq_results <- run_deseq_APP_Sex(Hippocampus_counts_subset_df, Hippocampus_meta_subset, 10, condition)
Hippocampus_results <- Hippocampus_deseq_results$res
Hippocampus_dds <- Hippocampus_deseq_results$dds
head(Hippocampus_results, 10)
Hippocampus_results
Hippocampus_dds
# write out DE results
write.csv(Hippocampus_results,'Hippocampus_deseq_results_APP_Sex.csv')
message("Hippocampus DESeq2 result is saved in file: 'Hippocampus_deseq_results_APP_Sex.csv' ")

# write out matrix of normalized counts
write.csv(counts(Hippocampus_dds,normalized=TRUE),'Hippocampus_deseq_norm_counts_APP_Sex.csv')
message("Hippocampus DESeq2 normalized counts is saved in file: 'Hippocampus_deseq_norm_counts_APP_Sex.csv' ")

#Add gene name to deseq results file and save the file
Hippocampus_deseq_res <- as.data.frame(Hippocampus_deseq_results$res)
Hippocampus_deseq_res_with_genename <- tibble::rownames_to_column(Hippocampus_deseq_res, "Geneid")
Hippocampus_deseq_res_with_genename$mgi_symbol <- new_Hippocampus_counts$mgi_symbol[match(Hippocampus_deseq_res_with_genename$Geneid, new_Hippocampus_counts$Geneid)]
Hippocampus_deseq_res_with_genename <- Hippocampus_deseq_res_with_genename %>% relocate(mgi_symbol, .after=Geneid)
write.csv(Hippocampus_deseq_res_with_genename, 'Hippocampus_deseq_results_with_genename_APP_Sex.csv')
message("Hippocampus DESeq2 result with Gene names (mgi_symbols) is saved in file: 'Hippocampus_deseq_results_with_genename_APP_Sex.csv' ")

############################## padj_threshold = 0.05 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.05
Hippocampus_deseq_results_padj_filtered <- filter(Hippocampus_deseq_res, padj < padj_threshold)
write.csv(Hippocampus_deseq_results_padj_filtered, 'Hippocampus_deseq_results_APP_Sex_padj_05_filtered.csv')
message("Hippocampus DESeq2 result after padj (0.05) filtering is saved in file: 'Hippocampus_deseq_results_APP_Sex_padj_05_filtered.csv' ")

# Number of genes significant at padj less than 0.05
message(paste0("Number of DE genes significant at padj < 0.05 for Hippocampus:APP (APP vs WT) while controlling Sex: ", nrow(Hippocampus_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Hippocampus_deseq_res) - nrow(Hippocampus_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold < 0.05 filtering for Hippocampus:APP (APP vs WT) while controlling Sex: ", no_of_genes_discarded))

Hippocampus_deseq_hist_title1 <- "Histogram of log2FoldChange:Hippocampus:APP (APP vs WT) - Sex Controlled:padj < 0.05"
Hippocampus_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_hist_title1)
Hippocampus_deseq_fc_hist_plot1

#scatter plot
Hippocampus_deseq_scat_title1 <- "Scatter Plot:Hippocampus:APP (APP vs WT) - Sex Controlled:padj < 0.05"
Hippocampus_deseq_scat_plot1 <- deseq_DE_scat_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_scat_title1)
Hippocampus_deseq_scat_plot1

#Add gene name to deseq results file and save the file
Hippocampus_deseq_padj_filtered_05 <- Hippocampus_deseq_results_padj_filtered
Hippocampus_deseq_padj_filtered_05 <- tibble::rownames_to_column(Hippocampus_deseq_padj_filtered_05, "Geneid")
Hippocampus_deseq_padj_filtered_05$mgi_symbol <- new_Hippocampus_counts$mgi_symbol[match(Hippocampus_deseq_padj_filtered_05$Geneid, new_Hippocampus_counts$Geneid)]
write.csv(Hippocampus_deseq_padj_filtered_05, 'Hippocampus_deseq_results_APP_Sex_padj_05_filtered_genename.csv')

############################## padj_threshold = 0.10 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.10
Hippocampus_deseq_results_padj_filtered <- filter(Hippocampus_deseq_res, padj < padj_threshold)
write.csv(Hippocampus_deseq_results_padj_filtered, 'Hippocampus_deseq_results_APP_Sex_padj_1_filtered.csv')
message("Hippocampus DESeq2 result after padj (0.1) filtering is saved in file: 'Hippocampus_deseq_results_APP_Sex_padj_1_filtered.csv' ")

# Number of genes significant at padj less than 0.1
message(paste0("Number of DE genes significant at padj < 0.1 for Hippocampus:APP (APP vs WT) while controlling Sex: ", nrow(Hippocampus_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Hippocampus_deseq_res) - nrow(Hippocampus_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold < 0.1 filtering for Hippocampus:APP (APP vs WT) while controlling Sex: ", no_of_genes_discarded))

Hippocampus_deseq_hist_title1 <- "Histogram of log2FoldChange:Cortex:APP (APP vs WT) - Sex Controlled:padj < 0.1"
Hippocampus_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_hist_title1)
Hippocampus_deseq_fc_hist_plot1

#scatter plot
Hippocampus_deseq_scat_title1 <- "Scatter Plot:Hippocampus:APP (APP vs WT) - Sex Controlled:padj < 0.1"
Hippocampus_deseq_scat_plot1 <- deseq_DE_scat_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_scat_title1)
Hippocampus_deseq_scat_plot1

#Add gene name to deseq results file and save the file
Hippocampus_deseq_padj_filtered_1 <- Hippocampus_deseq_results_padj_filtered
Hippocampus_deseq_padj_filtered_1 <- tibble::rownames_to_column(Hippocampus_deseq_padj_filtered_1, "Geneid")
Hippocampus_deseq_padj_filtered_1$mgi_symbol <- new_Hippocampus_counts$mgi_symbol[match(Hippocampus_deseq_padj_filtered_1$Geneid, new_Hippocampus_counts$Geneid)]
write.csv(Hippocampus_deseq_padj_filtered_1, 'Hippocampus_deseq_results_APP_Sex_padj_1_filtered_genename.csv')

```

10. Hippocampus: Enhanced Volcano Plot

```{r Part10}
# generate Enhanced Volcano Plot
padj_threshold <- 0.05
FC_threshold <- 0
subtitle <- "DESeq2 Results:Hippocampus:APP (APP vs WT) - Sex Controlled"
Hippocampus_EnhancedVolcano_plot <- plot_enhanced_volcano(Hippocampus_deseq_res_with_genename, padj_threshold,  FC_threshold, subtitle)
Hippocampus_EnhancedVolcano_plot

```

11. Hippocampus: Heatmap

```{r Part11}
 # generate heatmap
 Hippocampus_title <- "Heatmap:Hippocampus:APP (APP vs WT) - Sex Controlled"
 Hippocampus_results_with_genename <- Hippocampus_deseq_res_with_genename[, -c(1)]
 rownames(Hippocampus_results_with_genename) <- Hippocampus_deseq_res_with_genename$Geneid
 plot_heatmap <-  generate_heatmap(Hippocampus_dds, Hippocampus_results_with_genename, Hippocampus_title)
 plot_heatmap

```

12 Hippocampus: PCA

```{r Part12}

 # PCA
 message("************** Principal Component Analysis (PCA) **************")
 Hippocampus_ntd <- normTransform(Hippocampus_dds)
 Hippocampus_ntd
 #view(assay(Cortex_ntd))
 Hippocampus_pca_results <- prcomp(scale(t(assay(Hippocampus_ntd))), center=FALSE, scale=FALSE)
 summary(Hippocampus_pca_results)
 Hippocampus_pc_variance_explained <- calculate_variance_explained(Hippocampus_pca_results)
 Hippocampus_pc_variance_explained
 Hippocampus_variance_tibble <- make_variance_tibble(Hippocampus_pc_variance_explained, Hippocampus_pca_results)
 head(Hippocampus_variance_tibble, 5)
 plot_pca_variance(Hippocampus_variance_tibble)
 Hippocampus_biplot_title = "PCA Plot: PC1 vs PC2:Hippocampus:APP (APP vs WT) - Sex Controlled"
 make_biplot(Hippocampus_metadata_filename, Hippocampus_pca_results, Hippocampus_biplot_title)
 
``` 

13. GSEA:Running fgsea

Perform a GSEA using a ranked list log2FoldChange values for all genes discovered in the 
DESeq2 results against the M2 Canonical Pathways gene set collection. Convert gene identifiers in the results to the appropriate matching format found in the M2 gene sets. fgsea expects a named vector of gene level statistics and the gene sets in the form of a named list.

Using "m2.cp.wikipathways.v0.3.symbols.gmt" which is the Canonical Pathways gene sets 
derived from the WikiPathways pathway database (WikiPathways subset of CP)


```{r Part13}
#Label the results with the column as specified above
padj_threshold <- 0.05
FC_threshold <- 0
Cortex_labeled_results <- label_res(Cortex_results, padj_threshold, FC_threshold)
Cortex_labeled_results <- Cortex_labeled_results[order(Cortex_labeled_results$padj, decreasing = FALSE), ]
Cortex_labeled_results %>% relocate(genes, volc_plot_status, log2FoldChange, padj)
#Run fgsea using a ranked list of descending log2FC against the wiki pathways gene set
Cortex_fgsea_results <- run_gsea(Cortex_labeled_results, '../data/m2.cp.wikipathways.v0.3.symbols.gmt')
Cortex_filtered_results <- Cortex_fgsea_results[order(Cortex_fgsea_results$NES, decreasing = FALSE), ]
Cortex_filtered_results
Cortex_fgsea_results_mat <- as.matrix(Cortex_filtered_results)
write.csv(Cortex_fgsea_results_mat, 'Cortex_NES_ordered_fgsea_results_APP_Sex_FC0.csv')
message("Cortex NES ordered fgsea results saved in file: 'Cortex_NES_ordered_fgsea_results_APP_Sex_FC0.csv' ")

#Plot the top ten pathways with both positive and negative NES (20 total)
#Color the pathways by the sign of their NES (positive or negative)
Cortex_fgsea_title = "Cortex:APP (APP vs WT) - Sex Controlled"
Cortex_fgsea_plot <- top_pathways(Cortex_filtered_results, 10, Cortex_fgsea_title)
Cortex_fgsea_plot

#################################################################
# Hippocampus FGSEA 
FC_threshold <- 0
Hippocampus_labeled_results <- label_res(Hippocampus_results, padj_threshold, FC_threshold)
Hippocampus_labeled_results <- Hippocampus_labeled_results[order(Hippocampus_labeled_results$padj, decreasing = FALSE), ]
Hippocampus_labeled_results %>% relocate(genes, volc_plot_status, log2FoldChange, padj)
Hippocampus_fgsea_results <- run_gsea(Hippocampus_labeled_results, '../data/m2.cp.wikipathways.v0.3.symbols.gmt')
Hippocampus_filtered_results <- Hippocampus_fgsea_results[order(Hippocampus_fgsea_results$NES, decreasing = FALSE), ]
Hippocampus_filtered_results
Hippocampus_fgsea_results_mat <- as.matrix(Hippocampus_filtered_results)
write.csv(Hippocampus_fgsea_results_mat, 'Hippocampus_NES_ordered_fgsea_results_APP_Sex_FC0.csv')
message("Hippocampus NES ordered fgsea results saved in file: 'Hippocampus_NES_ordered_fgsea_results_APP_Sex_FC0.csv' ")

#Plot the top ten pathways with both positive and negative NES (20 total)
#Color the pathways by the sign of their NES (positive or negative)
#top_pathways(fgsea_results, 10)
Hippocampus_fgsea_title = "Hippocampus:APP (APP vs WT) - Sex Controlled"
Hippocampus_fgsea_plot <- top_pathways(Hippocampus_filtered_results, 10, Hippocampus_fgsea_title)
Hippocampus_fgsea_plot

```

