---
title: "Summer Internship"
author: "Navin Ramanan"
date: "6/01/2022"
output: html_document
---

##Setup
```{r Packages + setup, include=FALSE, results='hide'}
#Source your main
source('Diet_APP_9.R') # load all our wonderful functions
knitr::opts_chunk$set(echo = FALSE) # no markdown for you <3

```

1.  Map Ensembl ID  to Gene Symbol in the counts file

```{r Part1}
#Read in verse_counts.tsv and display the head of the resulting tibble
Hippocampus_counts_filename2 <- '../data/Hippocampus_counts_matrix2.csv'
Cortex_counts_filename2 <- '../data/Cortex_counts_matrix2.csv'

Hippocampus_counts2 <- read.csv(Hippocampus_counts_filename2,as.is=TRUE, check.names=FALSE)
Cortex_counts2 <- read.csv(Cortex_counts_filename2,as.is=TRUE, check.names=FALSE)

Hippocampus_gene_name2 <- ensembl_id_to_gene_symbol(Hippocampus_counts2$Geneid)
Cortex_gene_name2 <- ensembl_id_to_gene_symbol(Cortex_counts2$Geneid)

```

2.  Write Counts Data with Gene Symbol to CSV

```{r Part2}

Hippocampus_geneids <- Hippocampus_gene_name2
Cortex_geneids <- Cortex_gene_name2
names(Hippocampus_geneids)[names(Hippocampus_geneids) == "ensembl_gene_id"] <- "Geneid"
names(Cortex_geneids)[names(Cortex_geneids) == "ensembl_gene_id"] <- "Geneid"

#remove duplicate rows across specific columns of data frame
Hippocampus_Geneid <- Hippocampus_geneids %>% distinct(Geneid, .keep_all = TRUE)
Cortex_Geneid <- Cortex_geneids %>% distinct(Geneid, .keep_all = TRUE)

new_Hippocampus_counts <- merge(Hippocampus_geneids, Hippocampus_counts2, by='Geneid')
new_Cortex_counts <- merge(Cortex_geneids, Cortex_counts2, by='Geneid')

# write counts table with mice gene name to a csv file
#write.csv(new_Hippocampus_counts,'Hippocampus_counts_with_gene_names.csv')
#write.csv(new_Cortex_counts,'Cortex_counts_with_gene_names.csv')

```

3.  Reading and filtering Hippocampus count and metadata files

```{r Part3}
#Read in verse_counts.tsv and display the head of the resulting tibble
Hippocampus_counts_filename <- '../data/Hippocampus_counts_matrix.csv'
Hippocampus_metadata_filename <- '../data/Hippocampus_metadata.csv'

# metadata
Hippocampus_metadata <- read.csv(Hippocampus_metadata_filename,as.is=TRUE)
# Make the first col of metadata into rownames
Hippocampus_meta_subset <- subset(Hippocampus_metadata, select= -c(1))
rownames(Hippocampus_meta_subset) <- Hippocampus_metadata[,1]

# Count data
Hippocampus_counts <- read.csv(Hippocampus_counts_filename,as.is=TRUE, check.names=FALSE)
# Keep only Counts cols and make Geneid col into rownames
Hippocampus_counts_subset <- subset(Hippocampus_counts, select= -c(1,2,3,4,5,6))
rownames(Hippocampus_counts_subset) <- Hippocampus_counts[,1]

# Keep rows with sum of sample counts greater than 10
Hippocampus_counts_subset_df <- Hippocampus_counts_subset[(rowSums(Hippocampus_counts_subset) > 10),]

```

4.  Reading and filtering Cortex count and metadata files

```{r Part4}
#Read in verse_counts.tsv and display the head of the resulting tibble
Cortex_counts_filename <- '../data/Cortex_counts_matrix.csv'
Cortex_metadata_filename <- '../data/Cortex_metadata.csv'

# metadata
Cortex_metadata <- read.csv(Cortex_metadata_filename,as.is=TRUE)
# Make the first col of metadata into rownames
Cortex_meta_subset <- subset(Cortex_metadata, select= -c(1))
rownames(Cortex_meta_subset) <- Cortex_metadata[,1]

# Count data
Cortex_counts <- read.csv(Cortex_counts_filename,as.is=TRUE, check.names=FALSE)
# Keep only Counts cols and make Geneid col into rownames
Cortex_counts_subset <- subset(Cortex_counts, select= -c(1,2,3,4,5,6))
rownames(Cortex_counts_subset) <- Cortex_counts[,1]

# Keep rows with sum of sample counts greater than 10
Cortex_counts_subset_df <- Cortex_counts_subset[(rowSums(Cortex_counts_subset) > 10),]
#view(Cortex_counts_subset_df)

```

5.  Cortex: APP=APP, Age=9, Diet=Control vs Supplemented (~APP + Age + Diet)
            APP9Control vs APP9Supplemented
   
```{r Part5}

message("Cortex: Running DESeq2 for Diet(Control vs Supplemented) with APP=APP, Age=9:padj < 0.05 and 0.10 ")
message("Cortex: APP9Control vs APP9Supplemented")
condition = c("group", "APP9Control", "APP9Supplemented")
Cortex_deseq_results <- run_deseq_Diet_APP_Age(Cortex_counts_subset_df, Cortex_meta_subset, 10, condition)
Cortex_results <- Cortex_deseq_results$res
Cortex_dds <- Cortex_deseq_results$dds
#view(Cortex_results)
head(Cortex_results, 10)
Cortex_results
Cortex_dds
# write out DE results
write.csv(Cortex_results,'Cortex_deseq_results_Diet_APP_9.csv')
message("Cortex DESeq2 result is saved in file: 'Cortex_deseq_results_Diet_APP_9.csv' ")

# write out matrix of normalized counts
write.csv(counts(Cortex_dds,normalized=TRUE),'Cortex_deseq_norm_counts_Diet_APP_9.csv')
message("Cortex DESeq2 normalized counts is saved in file: 'Cortex_deseq_norm_counts_Diet_APP_9.csv' ")

#Add gene name to deseq results file and save the file
Cortex_deseq_res <- as.data.frame(Cortex_deseq_results$res)
Cortex_deseq_res_with_genename <- tibble::rownames_to_column(Cortex_deseq_res, "Geneid")
Cortex_deseq_res_with_genename$mgi_symbol <- new_Cortex_counts$mgi_symbol[match(Cortex_deseq_res_with_genename$Geneid, new_Cortex_counts$Geneid)]
Cortex_deseq_res_with_genename <- Cortex_deseq_res_with_genename %>% relocate(mgi_symbol, .after=Geneid)
write.csv(Cortex_deseq_res_with_genename, 'Cortex_deseq_results_with_genename_Diet_APP_9.csv')
message("Cortex DESeq2 result with Gene names (mgi_symbols) is saved in file: 'Cortex_deseq_results_with_genename_Diet_APP_9.csv' ")

############################## padj_threshold = 0.05 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.05
print(nrow(Cortex_deseq_res))
Cortex_deseq_results_padj_filtered <- filter(Cortex_deseq_res, padj < padj_threshold)
print(nrow(Cortex_deseq_results_padj_filtered))
write.csv(Cortex_deseq_results_padj_filtered, 'Cortex_deseq_results_Diet_APP_9_padj_05_filtered.csv')
message("Cortex DESeq2 result after padj (0.05) filtering is saved in file: 'Cortex_deseq_results_Diet_APP_9_padj_05_filtered.csv' ")

# Number of genes significant at padj less than 0.05
message(paste0("Number of DE genes significant at padj < 0.05 for Cortex:APP9Control vs APP9Supplemented: ", nrow(Cortex_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Cortex_deseq_res) - nrow(Cortex_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold 0.05 filtering for Cortex:APP9Control vs APP9Supplemented: ", no_of_genes_discarded))
  
#Histograms of fold change values from the significant DE genes. 
Cortex_deseq_hist_title1 <- "Histogram of log2FoldChange:Cortex:APP9Control vs APP9Supplemented:padj < 0.05"
Cortex_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_hist_title1)
Cortex_deseq_fc_hist_plot1

#Scatter plots of fold change vs nominal p-value.
Cortex_deseq_scat_title1 <- "Scatter Plot:Cortex:APP9Control vs APP9Supplemented:padj < 0.05"
Cortex_deseq_scat_plot1 <- deseq_DE_scat_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_scat_title1)
Cortex_deseq_scat_plot1

############################## padj_threshold = 0.10 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.10
Cortex_deseq_res <- as.data.frame(Cortex_deseq_results$res)
print(nrow(Cortex_deseq_res))
Cortex_deseq_results_padj_filtered <- filter(Cortex_deseq_res, padj < padj_threshold)
print(nrow(Cortex_deseq_results_padj_filtered))
write.csv(Cortex_deseq_results_padj_filtered, 'Cortex_deseq_results_Diet_APP_9_padj_1_filtered.csv')
message("Cortex DESeq2 result after padj (0.1) filtering is saved in file: 'Cortex_deseq_results_Diet_APP_9_padj_1_filtered.csv' ")

# Number of genes significant at padj less than 0.1
message(paste0("Number of DE genes significant at padj < 0.1 for Cortex:APP9Control vs APP9Supplemented: ", nrow(Cortex_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Cortex_deseq_res) - nrow(Cortex_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold 0.1 filtering for Cortex:APP9Control vs APP9Supplemented: ", no_of_genes_discarded))
  
#Histograms of fold change values from the significant DE genes. 
Cortex_deseq_hist_title1 <- "Histogram of log2FoldChange:Cortex:APP9Control vs APP9Supplemented:padj < 0.1"
Cortex_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_hist_title1)
Cortex_deseq_fc_hist_plot1

#Scatter plots of fold change vs nominal p-value.
Cortex_deseq_scat_title1 <- "Scatter Plot:Cortex:APP9Control vs APP9Supplemented:padj < 0.1"
Cortex_deseq_scat_plot1 <- deseq_DE_scat_plot(Cortex_deseq_results_padj_filtered, Cortex_deseq_scat_title1)
Cortex_deseq_scat_plot1

```
6. Cortex: Enhanced Volcano Plot

```{r Part6}
# generate Enhanced Volcano Plot
padj_threshold <- 0.05
FC_threshold <- 0
subtitle <- "DESeq2 Results:Cortex:APP9Control vs APP9Supplemented"
Cortex_EnhancedVolcano_plot <- plot_enhanced_volcano(Cortex_deseq_res_with_genename, padj_threshold,  FC_threshold, subtitle)
Cortex_EnhancedVolcano_plot

```

7. Cortex: Heatmap

```{r Part 7}
 # generate heatmap
 Cortex_title <- "Heatmap:Cortex:APP9Control vs APP9Supplemented"
 Cortex_results_with_genename <- Cortex_deseq_res_with_genename[, -c(1)]
 rownames(Cortex_results_with_genename) <- Cortex_deseq_res_with_genename$Geneid
 plot_heatmap <-  generate_heatmap(Cortex_dds, Cortex_results_with_genename, Cortex_title)
 plot_heatmap

```

8.  Hippocampus: APP=APP, Age=9, Diet=Control vs Supplemented (~APP + Age + Diet)
                 APP9Control vs APP9Supplemented

```{r Part8}
message("Hippocampus: Running DESeq2 for Diet(Control vs Supplemented) with APP=APP, Age=9:padj < 0.05 and 0.10 ")
message("Hippocampus: APP9Control vs APP9Supplemented")
condition = c("group", "APP9Control", "APP9Supplemented")
Hippocampus_deseq_results <- run_deseq_Diet_APP_Age(Hippocampus_counts_subset_df, Hippocampus_meta_subset, 10, condition)
Hippocampus_results <- Hippocampus_deseq_results$res
Hippocampus_dds <- Hippocampus_deseq_results$dds
#view(Hippocampus_results)
head(Hippocampus_results, 10)
Hippocampus_results
Hippocampus_dds
# write out DE results
write.csv(Hippocampus_results,'Hippocampus_deseq_results_Diet_APP_9.csv')
message("Hippocampus DESeq2 result is saved in file: 'Hippocampus_deseq_results_Diet_APP_9.csv' ")

# write out matrix of normalized counts
write.csv(counts(Hippocampus_dds,normalized=TRUE),'Hippocampus_deseq_norm_counts_Diet_APP_9.csv')
message("Hippocampus DESeq2 normalized counts is saved in file: 'Hippocampus_deseq_norm_counts_Diet_APP_9.csv' ")

#Add gene name to deseq results file and save the file
#view(Hippocampus_results)
Hippocampus_deseq_res <- as.data.frame(Hippocampus_deseq_results$res)
Hippocampus_deseq_res_with_genename <- tibble::rownames_to_column(Hippocampus_deseq_res, "Geneid")
Hippocampus_deseq_res_with_genename$mgi_symbol <- new_Hippocampus_counts$mgi_symbol[match(Hippocampus_deseq_res_with_genename$Geneid, new_Hippocampus_counts$Geneid)]
Hippocampus_deseq_res_with_genename <- Hippocampus_deseq_res_with_genename %>% relocate(mgi_symbol, .after=Geneid)
write.csv(Hippocampus_deseq_res_with_genename, 'Hippocampus_deseq_results_with_genename_Diet_APP_9.csv')
message("Hippocampus DESeq2 result with Gene names (mgi_symbols) is saved in file: 'Hippocampus_deseq_results_with_genename_Diet_APP_9.csv' ")
#view(Hippocampus_deseq_res_with_genename)

############################## padj_threshold = 0.05 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.05
Hippocampus_deseq_res <- as.data.frame(Hippocampus_deseq_results$res)

Hippocampus_deseq_results_padj_filtered <- filter(Hippocampus_deseq_res, padj < padj_threshold)
write.csv(Hippocampus_deseq_results_padj_filtered, 'Hippocampus_deseq_results_Diet_APP_9_padj_05_filtered.csv')
message("Hippocampus DESeq2 result after padj (0.05) filtering is saved in file: 'Hippocampus_deseq_results_Diet_APP_9_padj_05_filtered.csv' ")

# Number of genes significant at padj less than 0.05
message(paste0("Number of DE genes significant at padj < 0.05 for Hippocampus:APP9Control vs APP9Supplemented: ", nrow(Hippocampus_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Hippocampus_deseq_res) - nrow(Hippocampus_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold < 0.05 filtering for Hippocampus:APP9Control vs APP9Supplemented: ", no_of_genes_discarded))

Hippocampus_deseq_hist_title1 <- "Histogram of log2FoldChange:Hippocampus:APP9Control vs APP9Supplemented:padj < 0.05"
Hippocampus_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_hist_title1)
Hippocampus_deseq_fc_hist_plot1

#scatter plot
Hippocampus_deseq_scat_title1 <- "Scatter Plot:Hippocampus:APP9Control vs APP9Supplemented:padj < 0.05"
Hippocampus_deseq_scat_plot1 <- deseq_DE_scat_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_scat_title1)
Hippocampus_deseq_scat_plot1

############################## padj_threshold = 0.10 ##############################
# filter rows with padj < padj_threshold
padj_threshold <- 0.10
Hippocampus_deseq_results_padj_filtered <- filter(Hippocampus_deseq_res, padj < padj_threshold)
write.csv(Hippocampus_deseq_results_padj_filtered, 'Hippocampus_deseq_results_Diet_APP_9_padj_1_filtered.csv')
message("Hippocampus DESeq2 result after padj (0.1) filtering is saved in file: 'Hippocampus_deseq_results_Diet_APP_9_padj_1_filtered.csv' ")

# Number of genes significant at padj less than 0.1
message(paste0("Number of DE genes significant at padj < 0.1 for Hippocampus:APP9Control vs APP9Supplemented: ", nrow(Hippocampus_deseq_results_padj_filtered)))

# Number of DE Genes discarded
no_of_genes_discarded <- nrow(Hippocampus_deseq_res) - nrow(Hippocampus_deseq_results_padj_filtered)
message(paste0("Number of DE genes discarded after padj threshold < 0.1 filtering for Hippocampus:APP9Control vs APP9Supplemented: ", no_of_genes_discarded))

Hippocampus_deseq_hist_title1 <- "Histogram of log2FoldChange:Hippocampus:APP9Control vs APP9Supplemented:padj < 0.10"
Hippocampus_deseq_fc_hist_plot1 <- deseq_FC_hist_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_hist_title1)
Hippocampus_deseq_fc_hist_plot1

#scatter plot
Hippocampus_deseq_scat_title1 <- "Scatter Plot:Hippocampus:APP9Control vs APP9Supplemented:padj < 0.10"
Hippocampus_deseq_scat_plot1 <- deseq_DE_scat_plot(Hippocampus_deseq_results_padj_filtered, Hippocampus_deseq_scat_title1)
Hippocampus_deseq_scat_plot1

```

9. Hippocampus: Enhanced Volcano Plot

```{r Part9}
# generate Enhanced Volcano Plot
padj_threshold <- 0.05
FC_threshold <- 0
subtitle <- "DESeq2 Results:Hippocampus:APP9Control vs APP9Supplemented"
Hippocampus_EnhancedVolcano_plot <- plot_enhanced_volcano(Hippocampus_deseq_res_with_genename, padj_threshold,  FC_threshold, subtitle)
Hippocampus_EnhancedVolcano_plot

```
10. Hippocampus: Heatmap

```{r Part10}
 # generate heatmap
 Hippocampus_title <- "Heatmap:Hippocampus:APP9Control vs APP9Supplemented"
 Hippocampus_results_with_genename <- Hippocampus_deseq_res_with_genename[, -c(1)]
 rownames(Hippocampus_results_with_genename) <- Hippocampus_deseq_res_with_genename$Geneid
 plot_heatmap <-  generate_heatmap(Hippocampus_dds, Hippocampus_results_with_genename, Hippocampus_title)
 plot_heatmap

```

11. GSEA:Running fgsea

Perform a GSEA using a ranked list log2FoldChange values for all genes discovered in the 
DESeq2 results against the M2 Canonical Pathways gene set collection. Convert gene identifiers in the results to the appropriate matching format found in the M2 gene sets. fgsea expects a named vector of gene level statistics and the gene sets in the form of a named list.

Using "m2.cp.wikipathways.v0.3.symbols.gmt" which is the Canonical Pathways gene sets 
derived from the WikiPathways pathway database (WikiPathways subset of CP)


```{r Part11}
#Label the results with the column as specified above
padj_threshold <- 0.05
FC_threshold <- 0
Cortex_labeled_results <- label_res(Cortex_results, padj_threshold, FC_threshold)
Cortex_labeled_results <- Cortex_labeled_results[order(Cortex_labeled_results$padj, decreasing = FALSE), ]
Cortex_labeled_results %>% relocate(genes, volc_plot_status, log2FoldChange, padj)
#view(Cortex_labeled_results)
#Run fgsea using a ranked list of descending log2FC against the wiki pathways gene set
Cortex_fgsea_results <- run_gsea(Cortex_labeled_results, '../data/m2.cp.wikipathways.v0.3.symbols.gmt')
Cortex_filtered_results <- Cortex_fgsea_results[order(Cortex_fgsea_results$NES, decreasing = FALSE), ]
Cortex_filtered_results
#view(Cortex_filtered_results)
Cortex_fgsea_results_mat <- as.matrix(Cortex_filtered_results)
write.csv(Cortex_fgsea_results_mat, 'Cortex_NES_ordered_fgsea_results_Diet_APP_9_FC0.csv')
message("Cortex NES ordered fgsea results saved in file: 'Cortex_NES_ordered_fgsea_results_Diet_APP_9_FC0.csv' ")

#Plot the top ten pathways with both positive and negative NES (20 total)
#Color the pathways by the sign of their NES (positive or negative)
Cortex_fgsea_title = "Cortex:APP9Control vs APP9Supplemented"
Cortex_fgsea_plot <- top_pathways(Cortex_filtered_results, 10, Cortex_fgsea_title)
Cortex_fgsea_plot

#################################################################
# Hippocampus FGSEA 
FC_threshold <- 0
Hippocampus_labeled_results <- label_res(Hippocampus_results, padj_threshold, FC_threshold)
Hippocampus_labeled_results <- Hippocampus_labeled_results[order(Hippocampus_labeled_results$padj, decreasing = FALSE), ]
Hippocampus_labeled_results %>% relocate(genes, volc_plot_status, log2FoldChange, padj)
Hippocampus_fgsea_results <- run_gsea(Hippocampus_labeled_results, '../data/m2.cp.wikipathways.v0.3.symbols.gmt')
Hippocampus_filtered_results <- Hippocampus_fgsea_results[order(Hippocampus_fgsea_results$NES, decreasing = FALSE), ]
Hippocampus_filtered_results
Hippocampus_fgsea_results_mat <- as.matrix(Hippocampus_filtered_results)
write.csv(Hippocampus_fgsea_results_mat, 'Hippocampus_NES_ordered_fgsea_results_Diet_APP_9_FC0.csv')
message("Hippocampus NES ordered fgsea results saved in file: 'Hippocampus_NES_ordered_fgsea_results_Diet_APP_9_FC0.csv' ")

#Plot the top ten pathways with both positive and negative NES (20 total)
#Color the pathways by the sign of their NES (positive or negative)
#top_pathways(fgsea_results, 10)
Hippocampus_fgsea_title = "Hippocampus:APP9Control vs APP9Supplemented"
Hippocampus_fgsea_plot <- top_pathways(Hippocampus_filtered_results, 10, Hippocampus_fgsea_title)
Hippocampus_fgsea_plot

```

